\documentclass[12pt]{article}

\usepackage{a4wide,outlines}

\title{Mentor Handbook Brainstorm Proceedings}
\author{Quinton Westrich \\ {\normalsize \texttt{qgwestrich21@tntech.edu}}}


\begin{document}

\maketitle
This is what we discussed on Thursday, March 22, 2007. Some general ideas and discussions were:
\begin{itemize}
	\item We might want to employ "google groups" and keep our documents in a semi-forum style.
	\item We can pull stuff from the web, but we must CITE it! using both the URL and the date the info was pulled.
	\item These are drafts and are not set in stone.
	\item Include real journals throughout, esp. from "dark" states/stages such as Perry 5, depression, showing different types of thinking and individuality in ways of working through common issues.
	\item Keep in mind what is essential for \emph{incoming freshmen} as you are writing and in compiling this workbook.
	\item It would be cool to include quotes throughout, e.g. at the beginning of each chapter.
	\item Try to put exercises and journals in every chapter.
	\item Write introductions to each chapter? vs. all in one chapter (I don't understand what I just typed)
	\item Include art throughout, esp. symbols, pictures
	\item We may want to keep this an online, continually expanding handbook.
	\item Quinton is good at typesetting (writing books). Nathan is good at graphics.
	\item We may want to delay disbursing a handbook to incoming students. This is intended to prevent overwhelming and to promote dealing with uncertainty.
\end{itemize}


\section{Chapter/Topic Contents}

\begin{outline}[enumerate]
\1 MBTI
	\2 How to speak to people with different types
	\2 Include exercises for high intuitives to practice S
		\3 E.g. cross-stitching
		\3 Drawing exactly what you see (nonjudgemental imagery)
\1 What is Mentor
	\2 Nominations for Mentor (sophomores and freshmen)
	\2 How to find nominees when Connie's not here
	\2 Steering Committee: Senior, level 2, interview people- see they \emph{can} do the commitments (journaling, meditation, attend meetings)
	\2 Bilaws- force them to stay open, constitution-style ammendments
\1 Depression, "Abnormal Psychology" (=Normal Psychology in Mentor)
	\2 Include Journals
	\2 Practical Skills
	\2 Training to recognize depression, mania, suicide, bipolar, schizophrenia, borderline, eating disorders
	\2 Black Holes: Depression indicator worksheet
	\2 Alcoholism, drugs, altered states
	\2 Addictions
	\2 Other Big Sib training (e.g. exercise to find where to take a student with a problem)
\1 Models: Perry, etc.
	\2 Include Journals
	\2 Autonomy/Authorities
\1 Music
	\2 How it affects the brain (chemically, biologically)
	\2 Wisdom (?)
\1 Mentor Movies, Books, Music
\1 Values
	\2 Roger's Axioms
\1 Critical Thinking
	\2 Sophistry
	\2 Various "pseudo's" (pseudoscience, etc.)
\1 Meditation
	\2 Include work from Kornfield's \emph{A Path with Heart}
\1 Mysticism
\1 Wisdom Path
	\2 Practical Wisdom
		\3 Tact
\1 Self Help
	\2 Kairos Stuff
		\3 Mindtalk
		\3 Forgiveness
	\2 Relationships
	\2 Richo's stuff
		\3 Fear
	\2 Relationships
	\2 Boundaries
\1 The Big Picture
	\2 Philosophy
		\3 Brief History
		\3 Focus in epistemology
		\3 Focus in ontology
	\2 Science
		\3 Cosmology
		\3 Anthropology
		\3 Evolution
\1 Autobiography (could whoever suggested this elaborate on what he/she meant? -Q)
\1 Intellectualism (could whoever suggested this elaborate on what he/she meant? -Q)
\1 Intuition
	\2 Counterintuition
\1 Creativity/Mind Games
\1 Brain Chemistry/Consciousness
	\2 Perception of Time (read "as in time" beside Brain Research?)
\1 Values
	\2 Roger's Axioms
	\2 Ethics
	\2 Morality
\1 Diversity
	\2 Include Exercises
\1 History of Mentor
\1 Discipline/Commitments
\1 Discussion of Introductory Books/Questions
\1 Journaling/Speaking
\1 Progoff Weekend/Imagery
\1 Vocabulary/Jargon Glossary
\1 Religious Beliefs and Practices in Organized Religion
	\2 Fundamentalism
\1 Physical Fitness
	\2 Yoga
	\2 Tai Chi
	\2 Body-mind Connection
\1 Archetypes
\1 Sexuality
\1 What is mentor
	\2 What is growth?
\1 Awareness
\1 Politics
\1 Service
\1 Psychology
	\2 Maslow's Hierarchy of Needs
	\2 Gardner's Multiple Intelligences
\1 Social Random Interactions (could whoever suggested this elaborate on what he/she meant? -Q (hereafter denoted "?"))
\1 Nature/Botany
\1 Coolness (?)
\1 Mind Games
\1 Shadow
\1 \emph{The Four Quartets}
\1 \emph{Siddhartha}
\1 Humility
\1 A "Conclusion"??? (may be continually expanding)
\end{outline}

\section{Ideas for Exercises}

\begin{enumerate}
	\item assertiveness
	\item empathy
	\item johari
	\item nohari
	\item social interactions (?)
	\item relaxation (in meditation ch.)
	\item stress management techniques
\end{enumerate}
	
\begin{table}[!!!h]
\begin{tabular}{|l|l|}
	\hline
	Chapter/Topic & Writers \\
	\hline
	Music & Quinton, Eamon \\
	\hline
	Counterintuition & Quinton, Scott\\
	\hline
	Depression & Luc, Quinton \\
	\hline
	What is Mentor?\footnote{incl. bilaws, methods for selecting minnows} & Craig \\
	\hline
	Coolness & Scott \\
	\hline
	Perry, etc. & Craig, Luc, Avery, Quinton \\
	\hline
	Wisdom & Jon Jones, Connie \\
	\hline
	MBTI & Erik, Luc \\
	\hline
	Brain Chemistry & Quinton, Eamon \\
	\hline
	Hedonism & Eamon, Luc \\
	\hline
	Values (incl. Roger's Axioms) 
	& Quinton \\
	\hline
	Critical Thinking (incl. logical fallacies) & Quinton, Scott \\
	\hline
	Creativity/Mindgames& Quinton, Scott, Eamon, Avery, Connie\footnote{The notes are contradictory here. Creativity is listed twice with different people each time.} \\
	\hline
	Exercises & everybody \\
	\hline
	Jargon & (blank) \\
	\hline
	Meditation \& Mindfulness & Craig\\
	\hline
	Relationships & Eamon, Avery, Quinton \\
	\hline
	History of Mentor & Connie \\
	\hline
	What is growth? & Connie, Craig, Luc, Quinton \\
	\hline
	Introduction/Letter & Connie \\
	\hline
\end{tabular}
\caption{Writing Responsibilities}
\end{table}

At the end of the notes it is written "Be involved in a search." I don't know what this means or where to put it.

\end{document}
