\chapter{Personal Development (Connie)}

\begin{authors}
	Connie Hood
\end{authors}

\section{Introduction}

\section{Inventory of the Now Time}

\begin{center}THE HOOD-PROGOFF INVENTORY OF THE NOW TIME IN YOUR LIFE\end{center}

 Gather some notebook paper, a couple of pens, and if you have it,
 several sheets of unlined white paper. Plan to do this work in a quiet
 place where you can reflect and work undisturbed for an hour or so. Do
 not think hard about the questions. Write down what pops into your
 head. You can go back and think about your answers later. You can do
 the written parts of ths inventory on your computer if you are more
 comfortable doing that. Don't analyse and think about why or audience or
 anything else. Just respond to the questions. You can decide what to do
 with this exercise after you do it!

\medskip
 Anchor the Present
\medskip

 First let your mind wander back over over your life.... events and
 epochs in your life had their own beginnings and middles and ends...
 as your life flows into this present time in your life, sense the time
 you are in now....this now time in your life had its beginning but it
 has not come to an end.... it is an open time, with many
 possibilities....but many rivers run into this now time from your
 past.

 When you are ready, write out a focusing statement about this
 now time in your life, this time that has its beginning but is not yet
 finished.... for some people it will be coming to TTU; for others,
 connecting with special people; for others, making changes in
 vocational choices.... Begin the focusing statement with: ``It is a
 time when.....'' and finish the sentence.

\begin{enumerate}
	\item List some of the people who are important in this now time in your
 life.

 After you have listed them, go back and let images and feelings
 associated with each one of them come up in your mind. Write down one
 or two adjectives that you associate with each person. Did you list
 some people who have had 'negative' influences on your life? Who have
 hurt you or angered you? What adjectives can you place by their names?

	\item List projects that are important to you in this now time in your life.

	\item List events that have been important in this now time....

	\item Listen to your body.... What does it have to say to you about this now
 time in your life?

	\item What emotional states are dominant for you in this now time?

	\item What groups are important to you now...- university, social, church...?

	\item What dreams are recurring in this now time? or what dreams do you
 remember? just write a few words to remind yourself of what they
 are... don't try to analyze them.

	\item What are the most important \textbf{positive values} to you now?

	\item What are the most important \textbf{negative values} to you now? [Things you
 hate or dislike the most]

	\item Make a statement about your \textbf{religious} position now.

	\item What are the most \textbf{positive things} you can say about yourself now?

	\item What are the most \textbf{negative things} you can say about yourself now?

	\item What words come to mind that you associate with the \textbf{darkest part} of
 you, your shadow self?

	\item What image comes to mind?

	\item What strengths does your shadow have to offer?

	\item What are you most afraid of?

	\item What ecstasies have been most important to you? What peak experiences?

	\item What wisdom do you have to offer now? to the world? to yourself?

	\item What short term goals do you have for your future?

	\item What long term goals?

	\item What are your major concerns, issues, and problems in this NOW time?

	\item What is the payoff? What do you get out of it?

	\item Draw a picture of yourself with your non-dominant hand. [If you are
 right-handed, draw a picture of yourself with your left hand. Use
 plain white paper if you have it.]

	\item Draw a yen/yang symbol [ a circle with a capital S imbedded throughh
 the middle of it] and put in one side a symbol of your dark self and
 on the other side, a symbol of your light side.
\end{enumerate}

\medskip
 Fantasy Time
\medskip

\begin{enumerate}
\setcounter{enumi}{4}
	\item If there were no constraints, either moral or financial, what would
 you do with your life?

	\item What person that you don't know well or not at all, dead or alive,
 would you like to get to know?

	\item What project would you like to undertake now if there were no
 constraints?

	\item What event would you like to have happen now if you could have
 anything happen that you like?

	\item If you could build your ideal religion, what would you want as tenants
 or principles of it? would it be ideal for everyone? what would you
 like best about it?

	\item What would you like your body to be and do?

	\item What do you wish that you had learned?

	\item What would you like your strengths and talents to be?
\end{enumerate}

\section{Time Management}

\section{Neurosis and Psychosis}

\section{Codependency}

\section{Self Esteem}

\section{Boundaries and Assertiveness}

\section{Fear}

\section{Trust}

\section{Anger}

\section{Depression (Michael Roettger) and Suicide}

\section{Shadow (Connie)}

\section{Spirituality (Deanna and Connie)}

\section{Love and Unconditionality (Connie et al.}

\section{Creativity}


\section{Quotes}

\section{Works Cited}

\section{Further Reading}




\bibliographystyle{mla}		% the style you want to use for references.
\bibliography{mr,refs}				% the files containing all the articles and books you ever referenced.
