\documentclass[12pt,a4paper,twoside]{article}

% This file was converted to LaTeX by Writer2LaTeX ver. 0.4
% see http://www.hj-gym.dk/~hj/writer2latex for more info
\usepackage[ascii]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsfonts,textcomp}
\usepackage{color}
\usepackage{calc}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, pagecolor=blue, urlcolor=blue}
% List styles
\newcounter{saveenum}
\newcommand\liststyleWWviiiNumvii{%
\renewcommand\theenumi{\arabic{enumi}}
\renewcommand\theenumii{\alph{enumii}}
\renewcommand\theenumiii{\roman{enumiii}}
\renewcommand\theenumiv{\arabic{enumiv}}
\renewcommand\labelenumi{\theenumi.}
\renewcommand\labelenumii{\theenumii.}
\renewcommand\labelenumiii{\theenumiii.}
\renewcommand\labelenumiv{\theenumiv.}
}
% Pages styles (master pages)
\makeatletter
\newcommand\ps@Standard{%
\renewcommand\@oddhead{}%
\renewcommand\@evenhead{}%
\renewcommand\@oddfoot{}%
\renewcommand\@evenfoot{}%
 \setlength\paperwidth{8.5in}\setlength\paperheight{11in}\setlength\voffset{-1in}\setlength\hoffset{-1in}\setlength\topmargin{0.7874in}\setlength\headheight{12pt}\setlength\headsep{0cm}\setlength\footskip{12pt+0cm}\setlength\textheight{11in-0.7874in-0.7874in-0cm-12pt-0cm-12pt}\setlength\oddsidemargin{0.7874in}\setlength\textwidth{8.5in-0.7874in-0.7874in}
 \renewcommand\thepage{\arabic{page}}
 \setlength{\skip\footins}{0.0398in}\renewcommand\footnoterule{\vspace*{-0.0071in}\noindent\textcolor{black}{\rule{0.25\columnwidth}{0.0071in}}\vspace*{0.0398in}}
}
\makeatother
\pagestyle{Standard}
\begin{document}
\clearpage{\bfseries
Consent Form}

\bigskip

The following points have been explained to me:

\liststyleWWviiiNumvii
\begin{enumerate}
\item The reason for this research is to study the workload and situation awareness levels as a function of the number of robots being controlled.

\item The procedures are as follows:

\begin{enumerate}
\item The participant will be given a brief orientation.
\item The participant will be given a background questionnaire.
\item The participant will be trained on how to use the system.
\item The participant will complete two trials each having one task lasting 10 minutes.
\item After each task the participant will complete workload, situation awareness, and a brief questionnaire.
\item The participant will be given a brief final questionnaire after both trials.
\end{enumerate}

\item The user may discontinue the test at any point.

\item The discomforts or stresses that may be faced during this research are: None.

\item Participation entails the following risks: None.

\item The results of this participation will be anonymous and will not be released in any individually identifiable form without prior consent of the participant unless required by law.
\end{enumerate}

\bigskip

\line(1,0){300}

Signature of Investigator

\bigskip
\bigskip

\line(1,0){300}

Signature of Participant

\bigskip
\bigskip

\line(1,0){300}

Date

\bigskip

Please sign both copies. Keep one and return the other to the investigator.
\end{document}
